### 2023-03-19 Manuscript/Draft1.Rmd
small updates

<!-- [[I don't think we need to cite the trawl data because we said we were using trawl data above]] **BL: Does trawl data ever get cited? It was not cited above. cite trawl data?**. Trawl data is now cited above, suggest removing this comment. -->

@@

```{r paramsALL, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library("knitr")
library("kableExtra")
library("dplyr")
library("readr")
library("tidyr")
library("forcats")

@@

```{r paramsALL2,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library("knitr")
library("kableExtra")
library("dplyr")
library("readr")
library("tidyr")
library("forcats")

@@

<!-- ```{r,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'} -->
<!-- library("knitr") -->
<!-- library("kableExtra") -->
<!-- library("dplyr") -->
<!-- library("readr") -->
<!-- library("tidyr") -->
<!-- library("forcats") -->

@@

```{r paramsSCENARIOS,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library("knitr")
library("kableExtra")
library("dplyr")
library("readr")
library("tidyr")
library("forcats")

@@

<!-- ```{r YT_results,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'} -->
<!-- library("knitr") -->
<!-- library("kableExtra") -->
<!-- library("dplyr") -->
<!-- library("readr") -->
<!-- library("tidyr") -->
<!-- library("forcats") -->

@@

```{r YTresultsallstrata,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library("knitr")
library("kableExtra")
library("dplyr")
library("readr")
library("tidyr")
library("forcats")

@@

```{r YTresultsreducedstrata,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library("knitr")
library("kableExtra")
library("dplyr")
library("readr")
library("tidyr")
library("forcats")

@@

```{r CodERROR,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library("knitr")
library("kableExtra")
library("dplyr")
library("readr")
library("tidyr")
library("forcats")

@@

```{r HadERROR,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library("knitr")
library("kableExtra")
library("dplyr")
library("readr")
library("tidyr")
library("forcats")

>>>>

@@

```{r paramsALL, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}

@@

```{r paramsALL2,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}

@@

<!-- ```{r,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'} -->

@@

```{r paramsSCENARIOS,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}

@@

<!-- ```{r YT_results,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'} -->

@@

```{r YTresultsallstrata,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}

@@

```{r YTresultsreducedstrata,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}

@@

```{r CodERROR,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}

@@

```{r HadERROR,echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}

### 2023-03-19 Manuscript/Draft1.tex
small updates

much of below is from \url{https://apps-nefsc.fisheries.noaa.gov/nefsc/ecosystem-ecology/}
or
\url{https://www.fisheries.noaa.gov/data-tools/fisheries-economics-united-states-data-and-visualizations}

\begin{itemize}
\tightlist
\item
  The eastern continental shelf is ecologically diverse and economically important
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Bottom trawl survey is important for monitoring population trends
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Climate change is happening
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Briefly describe our study to test this
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Describe simulation study
\end{itemize}

@@

-- Used MixFishSim.

@@

Species-specific habitat preferences were derived using the \emph{lrren} tool from the R package \emph{envi} (Buller 2022) to create a niche model for each species. The \emph{lrren} tool estimates an ecological niche using the relative risk function by relating presence/absence data to two covariate predictors. We used bottom trawl point data in from 2009-2021 as our presence/absence input by using a value of 0 for any tow that failed to catch the given species and weighting a successful catch by the biomass of the given tow {[}{[}I don't think we need to cite the trawl data because we said we were using trawl data above{]}{]} \textbf{BL: Does trawl data ever get cited? It was not cited above. cite trawl data?}. We combined data from both the fall and spring surveys to obscure the influence of temperature so that the niche model would instead infer habitat preferences. Depth and mean sediment size were used as our covariate predictors. Estimated depth for the region was obtained from FVCOM (Chen et al. 2006). The mean sediment size raster was interpolated in ArcMap using the natural neighbor interpolation method \textbf{cite arcmap} using point data collected by the United States Geologic Survey (USGS) (McMullen et al. n.d.). Since the values in \(Hab^2_{J,s}\) are required to be between 0 and 1, we transform the spatial estimates from \emph{lrren} to fall between these bounds. See Figure \ref{fig:hab-plot1} for a visual representation of this process being applied to Cod. Figure \ref{fig:hab-plot} depicts habitat preferences \(Hab^2_{J,s}\) for each species.

@@

Each species is assumed to have normally distributed temperature preferences (\(N(\mu,\sigma)\)). We assume Yellowtail Flounder's preferences are \(N(8.75,4.25)\), while Haddock and Cod have preferences \(N(9,4)\). We chose these values by combining information in the literature with temperatures recorded in the bottom trawl survey. Weekly estimated temperature data for the region for 2012 was obtained from FVCOM (Chen et al. 2006). We chose to repeat estimates for a single year rather than use data for consecutive years to reduce the number of factors impacting model dynamics while still incorporating real data. The 2012 data was chosen because it displayed an average temperature pattern that consistently oscillated between maximum and minimum temperature values, allowing for a smooth repeating yearly temperature pattern for the constant temperature scenario. The 2012 temperature data was also transformed to create an oscillating pattern that increases 5 degrees Celsius on average over the duration of the simulation. We chose a 5 degree increase over a 20 year simulation to allow temperature change to have a meaningful impact on dynamics while remaining within reasonable computational limits in terms of the length of the simulation. Figure \ref{fig:temp-scenarios} depicts mean trends for the temperature scenarios used in our models. \textbf{dont forget to include gif in final submission}

@@

--Describe difference between increasing and constant temperature scenarios (images?)

@@

-- Describe each scenario that is considered

@@

-Describe post hoc sampling process and how data is used

@@

--Stratified mean vs VAST with and without covariates

@@

{[}{[} some things that I think we'll need to address, in addition to those mentioned above are (in no particular order):
number of simulations done for each scenario - time required to do them and analyses
pros and cons of using VAST with covariates in a changing environment
implications for current practice of using stratified mean (not horrible, but might be able to do better with model-based using covariates if enviro changing, note we don't want to say stratified mean should not be used because that will draw a lot of tomato throwing)
future work - some of the ideas we've discussed already could be mentioned
limitations to our study
perhaps for discussion, but we should mention somewhere that MFS only tracks biomass not age structure, the latter is important for stock assessments but not considered in this paper{]}{]}

>>>>

@@

@@

@@

@@

@@

@@

Species-specific habitat preferences were derived using the \emph{lrren} tool from the R package \emph{envi} (Buller 2022) to create a niche model for each species. The \emph{lrren} tool estimates an ecological niche using the relative risk function by relating presence/absence data to two covariate predictors. We used bottom trawl point data in from 2009-2021 as our presence/absence input by using a value of 0 for any tow that failed to catch the given species and weighting a successful catch by the biomass of the given tow We combined data from both the fall and spring surveys to obscure the influence of temperature so that the niche model would instead infer habitat preferences. Depth and mean sediment size were used as our covariate predictors. Estimated depth for the region was obtained from FVCOM (Chen et al. 2006). The mean sediment size raster was interpolated in ArcMap using the natural neighbor interpolation method \textbf{cite arcmap} using point data collected by the United States Geologic Survey (USGS) (McMullen et al. n.d.). Since the values in \(Hab^2_{J,s}\) are required to be between 0 and 1, we transform the spatial estimates from \emph{lrren} to fall between these bounds. See Figure \ref{fig:hab-plot1} for a visual representation of this process being applied to Cod. Figure \ref{fig:hab-plot} depicts habitat preferences \(Hab^2_{J,s}\) for each species.

@@

Each species is assumed to have normally distributed temperature preferences (\(N(\mu,\sigma)\)). We assume Yellowtail Flounder's preferences are \(N(8.75,4.25)\), while Haddock and Cod have preferences \(N(9,4)\). We chose these values by combining information in the literature with temperatures recorded in the bottom trawl survey. Weekly estimated temperature data for the region for 2012 was obtained from FVCOM (Chen et al. 2006). We chose to repeat estimates for a single year rather than use data for consecutive years to reduce the number of factors impacting model dynamics while still incorporating real data. The 2012 data was chosen because it displayed an average temperature pattern that consistently oscillated between maximum and minimum temperature values, allowing for a smooth repeating yearly temperature pattern for the constant temperature scenario. The 2012 temperature data was also transformed to create an oscillating pattern that increases 5 degrees Celsius on average over the duration of the simulation. We chose a 5 degree increase over a 20 year simulation to allow temperature change to have a meaningful impact on dynamics while remaining within reasonable computational limits in terms of the length of the simulation. Figure \ref{fig:temp-scenarios} depicts mean trends for the temperature scenarios used in our models. \textbf{dont forget to include animated gif in final submission}

@@

@@

@@

@@

@@
